---
params:
  ReportPark: "CHOH"
  ReportYear: 2019
  ReportAllYears: 2017-2021
output:
  html_document
---

<style>
.nav-pills>li>a{
padding: 8px;
font-size: 13px

}
</style>

```{r, include=F}
knitr::opts_chunk$set(echo=F)
knitr::opts_chunk$set(message=F)
knitr::opts_chunk$set(warning = F)


library(NPSForVeg)
library(tidyverse)
library(DT)
library(magrittr)

NCRN<-importNCRN("C:/Data/NCRN_Veg_2023/") #update data

Conditions<-read_csv("C:/Data/NCRN_Veg_2023/Tree_Sapling_Conditions.csv") #Update as well
names(NCRN)<-getNames(NCRN, name.class="code")
UsePark<-NCRN[[params$ReportPark]]
UsePlots<-getPlots(UsePark, years=params$ReportYear)$Plot_Name
AllPlots<-getPlots(UsePark)$Plot_Name
Commons<-getCommons(UsePark) %>% select(`Latin Name`=Latin_Name,`Common Name`=Common,Exotic) 

```

## `r params$ReportYear` Investigators Annual Report: `r getNames(NCRN[params$ReportPark], name.class ="long")` {.tabset .tabset-fade .tabset-pills} 

### Plot Map 

In `r paste0(params$ReportYear, ", ", length(UsePlots))` of `r length(AllPlots)` plots were monitored.  Monitoring data is summarized in the other tabs. The "`r paste(params$ReportYear,"Data")`" tab displays only the data collected in `r params$ReportYear`. The "`r paste("Full Park", min(params$ReportAllYears), "-", max(params$ReportAllYears), "Data")`" tab displays the most recent data for every plot monitored in entire park. The "New Species" tab indicates if any species were monitored for the first time in `r params$ReportYear`.

The monitoring protocol is available [here](https://irma.nps.gov/DataStore/Reference/Profile/2210263). 


On the map below, orange circles indicate plots monitored, black circles indicate plots that were not monitored.

```{r}
MapColors<-ifelse(AllPlots %in% UsePlots, 1, 2)

mapPlants(UsePark, plots=AllPlots, values=MapColors, colortype = "numeric", colors=c("orange", "black"))

```


###  `r paste(params$ReportYear,"Data")`{.tabset .tabset-fade .tabset-pills} 


#### Living Trees 

```{r, incldue=F}
if( getPlants(UsePark, "trees", years=params$ReportYear) %>% nrow() >0){
  TreeData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", values= "size", area="ha", Total=F) %>% select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(TreeData)<-c("Trees/ha", "Basal area (m2)/ha", "Number of Plots")
  TreeData<- TreeData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>%
    select(4,5,1,2,3,6)
} else TreeData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"trees", years=params$ReportYear) ))` individual living trees were monitored. `r if(!anyNA(TreeData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(TreeData)) datatable(TreeData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```



#### Dead Trees

```{r}
if( getPlants(UsePark, "trees", years=params$ReportYear, status="dead") %>% nrow() >0){
  DeadTreeData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", status="dead", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% 
      colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", status="dead", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(DeadTreeData)<-c("Trees/ha", "Number of Plots")

  DeadTreeData<- DeadTreeData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(3,4,1,2,5)
} else DeadTreeData<-NA

```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"trees", status="dead", years=params$ReportYear) ))` individual dead trees were monitored. `r if(!anyNA(DeadTreeData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(DeadTreeData)) datatable(DeadTreeData, options=list(pageLength=25)) %>% 
    formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))

```




#### Saplings

```{r}
if( getPlants(UsePark, "saplings", years=params$ReportYear) %>% nrow() >0){
  SapData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="saplings", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% 
      colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="saplings", values= "size", area="ha", Total=F) %>% select(-Plot_Name) %>% 
      colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="saplings", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(SapData)<-c("Saplings/ha", "Basal area (m2)/ha", "Number of Plots")
  SapData<- SapData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(4,5,1,2,3,6)
} else SapData<- NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"saplings", years=params$ReportYear) ))` individual living saplings were monitored. `r if(!anyNA(SapData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(SapData)) datatable(SapData, rownames=F, options=list(pageLength=25)) %>% 
      formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```




#### Tree Seedlings

```{r}
if( getPlants(UsePark, "seedlings", years=params$ReportYear) %>% nrow() >0){
  SeedData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="seedlings", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% 
      colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="seedlings", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(SeedData)<-c("Tree Seedlings/ha", "Number of Plots")
  SeedData<- SeedData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(3,4,1,2,5)
} else SeedData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"seedlings", years=params$ReportYear) ))` individual living tree seedlings were monitored. `r if(!anyNA(SeedData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(SeedData)) datatable(SeedData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```



#### Shrubs

```{r}
if( getPlants(UsePark, "shrubs", years=params$ReportYear) %>% nrow() >0){
  ShrubData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="shrubs", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="shrubs", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(ShrubData)<-c("Shrubs/ha", "Number of Plots")
  ShrubData<- ShrubData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>%  
    select(3,4,1,2,5)  
} else {ShrubData<-NA}
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"shrubs", years=params$ReportYear) ))` individual living shrubs were monitored. `r if(!anyNA(ShrubData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(ShrubData)) datatable(ShrubData, rownames=F, options=list(pageLength=25)) %>% 
    formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```



#### Shrub Seedlings

```{r}
if( getPlants(UsePark, "shseedlings", years=params$ReportYear) %>% nrow() >0){
  ShSeedData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="shseedlings", values="count", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="shseedlings", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(ShSeedData)<-c("Shrub Seedlings/ha", "Number of Plots")
  ShSeedData<- ShSeedData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>%  
    select(3,4,1,2,5)
} else ShSeedData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"shseedlings", years=params$ReportYear) ))` individual living shrub seedlings were monitored. `r if(!anyNA(ShSeedData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(ShSeedData)) datatable(ShSeedData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```



#### Vines on Trees

```{r}
if( getPlants(UsePark, "vines", years=params$ReportYear) %>% nrow() >0){
  VineData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="vines", values="count", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="vines", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(VineData)<-c("Vines/ha", "Number of Plots")
  VineData<- VineData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(3,4,1,2,5)
} else VineData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"vines", years=params$ReportYear) ))` individual living vines on trees were monitored. `r if(!anyNA(VineData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(VineData)) datatable(VineData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```



#### Understory Plants

```{r}
if( getPlants(UsePark, "herbs", years=params$ReportYear) %>% nrow() >0){
  HerbData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="herbs", values="size", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% divide_by(12) %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="herbs", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(HerbData)<-c("% Cover", "Number of Plots")
  HerbData<- HerbData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>%  left_join(Commons) %>% 
    select(3,4,1,2,5)
} else HerbData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"herbs", years=params$ReportYear) ))` observations of understory plants were recorded. `r if(!anyNA(HerbData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(HerbData)) datatable(HerbData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```


#### Pests and Diseases

```{r}
if( getPlants(UsePark, "trees", years=params$ReportYear) %>% nrow() >0 | getPlants(UsePark, "trees", years=params$ReportYear) %>% nrow() >0){
  PestData<-Conditions %>% filter(Unit_Code == params$ReportPark, Sample_Year==params$ReportYear, Pest==1 | Condition=="Beechleaf disease",
                                  Condition!="Other significant insect damage",
                                Status %in% c("Alive Broken","Alive Fallen", "Alive Leaning", "Alive Standing")) %>% 
    count(Condition)
    colnames(PestData)<-c("Pest or Disease", "Number of Occurances")
} else PestData<-NA

if( nrow(PestData)==0) PestData<-NA
```


In `r paste0(params$ReportYear, ", ", if(!is.data.frame(PestData)) "0" else nrow(PestData) )`, different pests or diseases were recorded on trees and saplings. `r if(!anyNA(PestData)) paste("See the table below for summary information on each species encountered.")`


```{r}
if (is.data.frame(PestData)) datatable(PestData, rownames=F, options=list(pageLength=25))
```


###  `r paste("Full Park", min(params$ReportAllYears), "-", max(params$ReportAllYears), "Data")` {.tabset .tabset-fade .tabset-pills} 


#### Living Trees 

```{r, incldue=F}
if( getPlants(UsePark, "trees", years=params$ReportAllYears) %>% nrow() >0){
  AllTreeData<-cbind(
    SiteXSpec(UsePark, years=params$ReportAllYears, group="trees", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="trees", values= "size", area="ha", Total=F) %>% select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="trees", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

   colnames(AllTreeData)<-c("Trees/ha", "Basal area (m2)/ha", "Number of Plots")
   AllTreeData <- AllTreeData %>% as.data.frame() %>%  mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>%  
   # AllTreeData<-AllTreeData %>% rowwise() %>% 
   #   mutate(`Common Name`=tryCatch(getPlantNames(UsePark, `Latin Name`), error=function(cond){" "})) %>% # this tryCatch business is due to species missing from the commonnames table (Rosaceae Family) 
     select(4,5,1,2,3,6)
} else AllTreeData<-NA
```


In `r paste( min(params$ReportAllYears),"-",max(params$ReportAllYears), ", ", nrow(getPlants(UsePark,"trees", years=params$ReportAllYears) ))` indvidual living trees were monitored. `r if(!anyNA(AllTreeData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(TreeData)) datatable(AllTreeData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```



#### Dead Trees

```{r}
if( getPlants(UsePark, "trees", years=params$ReportAllYears, status="dead") %>% nrow() >0){
  AllDeadTreeData<-cbind(
    SiteXSpec(UsePark, years=params$ReportAllYears, group="trees", status="dead", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% 
      colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="trees", status="dead", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(AllDeadTreeData)<-c("Trees/ha", "Number of Plots")

  AllDeadTreeData<- AllDeadTreeData %>% as.data.frame() %>% 
    mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(3,4,1,2,5)
} else AllDeadTreeData<-NA

```

In `r paste0(min(params$ReportAllYears),"-",max(params$ReportAllYears), ", ", nrow(getPlants(UsePark,"trees", status="dead", years=params$ReportAllYears) ))` individual dead trees were monitored. `r if(!anyNA(AllDeadTreeData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`




```{r}
if (is.data.frame(DeadTreeData)) datatable(DeadTreeData, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```




#### Saplings

```{r}
if( getPlants(UsePark, "saplings", years=params$ReportAllYears) %>% nrow() >0){
  AllSapData<-cbind(
    SiteXSpec(UsePark, years=params$ReportAllYears, group="saplings", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% 
      colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="saplings", values= "size", area="ha", Total=F) %>% select(-Plot_Name) %>% 
      colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="saplings", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(AllSapData)<-c("Saplings/ha", "Basal area (m2)/ha", "Number of Plots")
  AllSapData<- AllSapData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(4,5,1,2,3,6)
} else AllSapData<- NA
```


In `r paste0(min(params$ReportAllYears),"-",max(params$ReportAllYears), ", ", nrow(getPlants(UsePark,"saplings", years=params$ReportAllYears) ))` individual living saplings were monitored. `r if(!anyNA(SapData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(AllSapData)) datatable(AllSapData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```



#### Tree Seedlings

```{r}
if( getPlants(UsePark, "seedlings", years=params$ReportAllYears) %>% nrow() >0){
  AllSeedData<-cbind(
    SiteXSpec(UsePark, years=params$ReportAllYears, group="seedlings", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% 
      colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="seedlings", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(AllSeedData)<-c("Tree Seedlings/ha", "Number of Plots")
  AllSeedData<- AllSeedData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(3,4,1,2,5)
} else AllSeedData<-NA
```


In `r paste0(min(params$ReportAllYears),"-",max(params$ReportAllYears),", ", nrow(getPlants(UsePark,"seedlings", years=params$ReportAllYears) ))` individual living tree seedlings were monitored. `r if(!anyNA(AllSeedData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(AllSeedData)) datatable(AllSeedData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```



#### Shrubs

```{r}
if( getPlants(UsePark, "shrubs", years=params$ReportAllYears) %>% nrow() >0){
  AllShrubData<-cbind(
    SiteXSpec(UsePark, years=params$ReportAllYears, group="shrubs", values="count", area="ha", Total=F) %>% select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="shrubs", values= "presab", Total=F) %>% select(-Plot_Name) %>% colSums()
  )

  colnames(AllShrubData)<-c("Shrubs/ha", "Number of Plots")
  AllShrubData<- AllShrubData %>% as.data.frame() %>%    mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>%  
    select(3,4,1,2,5)  
} else {AllShrubData<-NA}
```


In `r paste0( min(params$ReportAllYears),"-",max(params$ReportAllYears), ", ", nrow(getPlants(UsePark,"shrubs", years=params$ReportAllYears) ))` individual living shrubs were monitored. `r if(!anyNA(AllShrubData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(AllShrubData)) datatable(AllShrubData, rownames=F, options=list(pageLength=25)) %>% 
   formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```


#### Shrub Seedlings

```{r}
if( getPlants(UsePark, "shseedlings", years=params$ReportAllYears) %>% nrow() >0){
  AllShSeedData<-cbind(
    SiteXSpec(UsePark, years=params$ReportAllYears, group="shseedlings", values="count", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="shseedlings", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(AllShSeedData)<-c("Shrub Seedlings/ha", "Number of Plots")
  AllShSeedData<- AllShSeedData %>% as.data.frame() %>% 
    mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>%
    select(3,4,1,2,5)
} else AllShSeedData<-NA
```


In `r paste0( min(params$ReportAllYears),"-",max(params$ReportAllYears), ", ", nrow(getPlants(UsePark,"shseedlings", years=params$ReportAllYears) ))` individual living shrub seedlings were monitored. `r if(!anyNA(AllShSeedData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (is.data.frame(AllShSeedData)) datatable(AllShSeedData, rownames=F, options=list(pageLength=25)) %>% 
   formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```




#### Vines on Trees

```{r}
if( getPlants(UsePark, "vines", years=params$ReportAllYears) %>% nrow() >0){
  AllVineData<-cbind(
    SiteXSpec(UsePark, years=params$ReportAllYears, group="vines", values="count", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="vines", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(AllVineData)<-c("Vines/ha", "Number of Plots")
  AllVineData<- AllVineData %>% as.data.frame() %>% 
    mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(3,4,1,2,5 )
} else AllVineData<-NA
```


In `r paste0( min(params$ReportAllYears),"-",max(params$ReportAllYears), ", ", nrow(getPlants(UsePark,"vines", years=params$ReportAllYears) ))` individual living vines on trees were monitored. `r if(!anyNA(AllVineData)) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (!anyNA(AllVineData)) datatable(AllVineData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```


#### Understory Plants

```{r}
if( getPlants(UsePark, "herbs", years=params$ReportAllYears) %>% nrow() >0){
  AllHerbData<-cbind(
    SiteXSpec(UsePark, years=params$ReportAllYears, group="herbs", values="size", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% divide_by(12) %>% round(1),
    SiteXSpec(UsePark, years=params$ReportAllYears, group="herbs", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(AllHerbData)<-c("% Cover", "Number of Plots")
  AllHerbData<- AllHerbData %>% as.data.frame() %>% 
    mutate(`Latin Name`= rownames(.)) %>% left_join(Commons) %>% 
    select(3,4,1,2,5)
} else AllHerbData<-NA
```


In `r paste0( min(params$ReportAllYears),"-",max(params$ReportAllYears),", ", nrow(getPlants(UsePark,"herbs", years=params$ReportAllYears) ))` observations of understory plants were recorded. `r if(!all(is.na(AllHerbData))) paste("See the table below for summary information on each species encountered. Rows that are shaded and have an Exotic value equal to 1, are exotic species.")`


```{r}
if (!all(is.na(AllHerbData))) datatable(AllHerbData, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))
```

#### Pests and Diseases

```{r}
if( getPlants(UsePark, "trees", years=params$ReportAllYears) %>% nrow() >0 | getPlants(UsePark, "trees", years=params$ReportAllYears) %>%
    nrow() >0){
    AllPestData<-Conditions %>% filter(Unit_Code == params$ReportPark, Sample_Year %in% params$ReportAllYears, Pest==1,
                                       Condition!="Other significant insect damage",
                                Status %in% c("Alive Broken","Alive Fallen", "Alive Leaning", "Alive Standing")) %>% 
    count(Condition)
    colnames(AllPestData)<-c("Pest or Disease", "Number of Occurances")
} else AllPestData<-NA

if( nrow(AllPestData)==0) AllPestData<-NA
```


In `r paste0(min(params$ReportAllYears),"-",max(params$ReportAllYears), ", ", nrow(AllPestData) )` different pests or diseases were recorded on trees and saplings. `r if(!all(is.na(AllPestData))) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(AllPestData)) datatable(AllPestData, rownames=F, options=list(pageLength=25))
```

###  `r paste("New Plant Species Found in", params$ReportYear)` 

```{r}
NewSpeciesTable<-function(park, group, reportyear, firstyear, commons_data){
  Recent_Speices<-getPlants(park, group, years=reportyear) %>% select(Latin_Name)
  Old_Species<-getPlants(park,group, years=firstyear:reportyear-1) %>% select(Latin_Name)
  New_Species<-anti_join(Recent_Speices,Old_Species) %>% distinct %>% 
    rename(`Latin Name`= Latin_Name) %>% left_join(Commons) %>% select(`Latin Name`, `Common Name`, Exotic)
  
return(list(Rows=nrow(New_Species), 
            Table=datatable(New_Species, rownames=F, options=list(pageLength=25)) %>% 
  formatStyle(columns="Exotic", target="row", backgroundColor = styleEqual(c(1,0), c("pink","white")))))
  
}

New_Trees<-NewSpeciesTable(UsePark, "trees", params$ReportYear, 2006, Commons)

New_Saps<-NewSpeciesTable(UsePark, "saplings", params$ReportYear, 2006, Commons)

New_Tree_Seeds<-NewSpeciesTable(UsePark, "seedlings", params$ReportYear, 2006, Commons)

New_Shrubs<-NewSpeciesTable(UsePark, "shrubs", params$ReportYear, 2006, Commons)

New_Shrub_Seeds<-NewSpeciesTable(UsePark, "shseedlings", params$ReportYear, 2006, Commons)

New_Vines<-NewSpeciesTable(UsePark, "vines", params$ReportYear, 2006, Commons)

New_Herbs<-NewSpeciesTable(UsePark, "herbs", params$ReportYear, 2006, Commons)


```

#### Trees

 
`r if(New_Trees$Rows>0) New_Trees$Table else paste("No new tree species found in ", params$ReportYear)`

<br>

#### Saplings

`r if(New_Saps$Rows>0) New_Saps$Table else paste("No new sapling species found in ", params$ReportYear)`

<br>

#### Tree Seedlings

`r if(New_Tree_Seeds$Rows>0) New_Tree_Seeds$Table else paste("No new tree seedlings species found in ", params$ReportYear)`

<br>

#### Shrubs

`r if(New_Shrubs$Rows>0) New_Shrubs$Tabe else paste("No new shrub species found in ", params$ReportYear)`

<br>

#### Shrub Seedlings

`r if(New_Shrub_Seeds$Rows>0) New_Shrub_Seeds$Table else paste("No new shrub seedling species found in ", params$ReportYear)`

<br>

#### Vines on Trees

`r if(New_Vines$Rows>0) New_Vines$Table else paste("No new vine species found in ", params$ReportYear)`

<br>

#### Understory Plants

`r if(New_Herbs$Rows>0) New_Herbs$Table else paste("No new understory plant species found in ", params$ReportYear)`




