---
params:
  ReportPark: "WOTR"
  ReportYear: 2019
output:
  html_document
---

```{r, include=F}
knitr::opts_chunk$set(echo=F)
knitr::opts_chunk$set(message=F)
knitr::opts_chunk$set(warning = F)


library(NPSForVeg)
library(tidyverse)
library(DT)
library(magrittr)

NCRN<-importNCRN("T:/I&M/MONITORING/FOREST_VEGETATION/08_DATA/2019/Exports/20191021") #update data
names(NCRN)<-getNames(NCRN, name.class="code")
UsePark<-NCRN[[params$ReportPark]]
UsePlots<-getPlots(UsePark, years=params$ReportYear)$Plot_Name

```

## `r params$ReportYear` Investigators Annual Report: `r getNames(NCRN[params$ReportPark], name.class ="long")` {.tabset .tabset-fade .tabset-pills} 

### Plots 

In `r paste0(params$ReportYear, ", ", length(UsePlots))`, plots were monitored. Refer to the map below for their location.

```{r}

mapPlants(UsePark, plots=UsePlots, values=1, colortype = "numeric", colors="darkgreen")

```


### Living Trees


```{r, incldue=F}
if( getPlants(UsePark, "trees", years=params$ReportYear) %>% nrow() >0){
  TreeData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", values="count", area="ha", Total=F)[,-1] %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", values= "size", area="ha", Total=F)[,-1] %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", values= "presab", Total=F)[,-1] %>% colSums()
  )

  colnames(TreeData)<-c("Trees/ha", "Basal area (m2)/ha", "Number of Plots")
  TreeData<- TreeData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.), `Common Name`=getPlantNames(UsePark, `Latin Name`)) %>% 
    select(4,5,1,2,3)
} else TreeData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"trees", years=params$ReportYear) ))` indvidual living trees were monitored. `r if(!anyNA(TreeData)) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(TreeData)) datatable(TreeData, rownames=F, options=list(pageLength=25))
```


### Dead Trees


```{r}
if( getPlants(UsePark, "trees", years=params$ReportYear, status="dead") %>% nrow() >0){
  DeadTreeData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", status="dead", values="count", area="ha", Total=F)[,-1] %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="trees", status="dead", values= "presab", Total=F)[,-1] %>% colSums()
  )

  colnames(DeadTreeData)<-c("Trees/ha", "Number of Plots")

  DeadTreeData<- DeadTreeData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.), `Common Name`=getPlantNames(UsePark, `Latin Name`)) %>% 
    select(3,4,1,2)
} else DeadTreeData<-NA

```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"trees", status="dead", years=params$ReportYear) ))` indvidual dead trees were monitored. `r if(!anyNA(DeadTreeData)) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(DeadTreeData)) datatable(DeadTreeData, options=list(pageLength=25))
```




### Saplings
```{r}
if( getPlants(UsePark, "saplings", years=params$ReportYear) %>% nrow() >0){
  SapData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="saplings", values="count", area="ha", Total=F)[,-1] %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="saplings", values= "size", area="ha", Total=F)[,-1] %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="saplings", values= "presab", Total=F)[,-1] %>% colSums()
  )

  colnames(SapData)<-c("Saplings/ha", "Basal area (m2)/ha", "Number of Plots")
  SapData<- SapData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.), `Common Name`=getPlantNames(UsePark, `Latin Name`)) %>% 
    select(4,5,1,2,3)
} else SapData<- NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"saplings", years=params$ReportYear) ))` indvidual living saplings were monitored. `r if(!anyNA(SapData)) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(SapData)) datatable(SapData, rownames=F, options=list(pageLength=25))
```





### Tree Seedlings

```{r}
if( getPlants(UsePark, "seedlings", years=params$ReportYear) %>% nrow() >0){
  SeedData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="seedlings", values="count", area="ha", Total=F)[,-1] %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="seedlings", values= "presab", Total=F)[,-1] %>% colSums()
  )

  colnames(SeedData)<-c("Tree Seedlings/ha", "Number of Plots")
  SeedData<- SeedData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.), `Common Name`=getPlantNames(UsePark, `Latin Name`)) %>% 
    select(3,4,1,2)
} else SeedData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"seedlings", years=params$ReportYear) ))` indvidual living tree seedlings were monitored. `r if(!anyNA(SeedData)) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(SeedData)) datatable(SeedData, rownames=F, options=list(pageLength=25))
```



### Shrubs

```{r}
if( getPlants(UsePark, "shrubs", years=params$ReportYear) %>% nrow() >0){
  ShrubData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="shrubs", values="count", area="ha", Total=F)[,-1] %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="shrubs", values= "presab", Total=F)[,-1] %>% colSums()
  )

  colnames(ShrubData)<-c("Shrubs/ha", "Number of Plots")
  ShrubData<- ShrubData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.), `Common Name`=getPlantNames(UsePark, `Latin Name`)) %>% 
    select(3,4,1,2)  
} else {ShrubData<-NA}
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"shrubs", years=params$ReportYear) ))` individual living shrubs were monitored. `r if(!anyNA(ShrubData)) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(ShrubData)) datatable(ShrubData, rownames=F, options=list(pageLength=25))
```

### Shrub Seedlings

```{r}
if( getPlants(UsePark, "shseedlings", years=params$ReportYear) %>% nrow() >0){
  ShSeedData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="shseedlings", values="count", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="shseedlings", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(ShSeedData)<-c("Shrub Seedlings/ha", "Number of Plots")
  ShSeedData<- ShSeedData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.), `Common Name`=getPlantNames(UsePark, `Latin Name`)) %>% 
    select(3,4,1,2)
} else ShSeedData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"shseedlings", years=params$ReportYear) ))` indvidual living shrub seedlings were monitored. `r if(!anyNA(ShSeedData)) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(ShSeedData)) datatable(ShSeedData, rownames=F, options=list(pageLength=25))
```

### Vines on Trees

```{r}
if( getPlants(UsePark, "vines", years=params$ReportYear) %>% nrow() >0){
  VineData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="vines", values="count", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="vines", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(VineData)<-c("Vines/ha", "Number of Plots")
  VineData<- VineData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.), `Common Name`=getPlantNames(UsePark, `Latin Name`)) %>% 
    select(3,4,1,2)
} else VineData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"vines", years=params$ReportYear) ))` indvidual living vines on trees were monitored. `r if(!anyNA(VineData)) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(VineData)) datatable(VineData, rownames=F, options=list(pageLength=25))
```


### Understory Plants

```{r}
if( getPlants(UsePark, "herbs", years=params$ReportYear) %>% nrow() >0){
  HerbData<-cbind(
    SiteXSpec(UsePark, years=params$ReportYear, group="herbs", values="size", area="ha", Total=F) %>% 
      select(-Plot_Name) %>% colSums() %>% divide_by(12) %>% round(1),
    SiteXSpec(UsePark, years=params$ReportYear, group="herbs", values= "presab", Total=F) %>% 
      select(-Plot_Name) %>% colSums()
  )

  colnames(HerbData)<-c("% Cover", "Number of Plots")
  HerbData<- HerbData %>% as.data.frame() %>% mutate(`Latin Name`= rownames(.), `Common Name`=getPlantNames(UsePark, `Latin Name`)) %>% 
    select(3,4,1,2)
} else HerbData<-NA
```


In `r paste0(params$ReportYear, ", ", nrow(getPlants(UsePark,"herbs", years=params$ReportYear) ))` observations of understory plants were recorded. `r if(!anyNA(HerbData)) paste("See the table below for summary information on each species encountered")`


```{r}
if (!anyNA(HerbData)) datatable(HerbData, rownames=F, options=list(pageLength=25))
```